Procrastination as a Marker of Cognitive Decline: Evidence from Longitudinal Transitions in the Older Adult Population

### 2025-11-07 manuscript/manuscript.qmd
Using new shifu template

This property asserts that the probability of transitioning from state \$X_t =
i\$ to a future state \$X\_{t+1} = j\$ depends only on the current state
\$X_t\$, and not on the full history of preceding states. Here, there are three
potential states: normative cognitive function, mild cognitive impairment, and
dementia. We consider dementia to be an absorbing state, i.e. once an individual
reaches this state, they cannot return to the other states in a future time
point.

@@

# Table information
tbl_descriptives <- tibble::tribble(
  ~"", ~Full, ~NC, ~MCI, ~Dementia,
  "Age (years)", "69.70 ± 7.58", "69.70 ± 7.55",	"69.60 ± 7.55",	"70.00 ± 9.89",
  "Male", "38.43% (n = 211)", "39.82% (n =180)", "32.56% (n = 28)", "27.27% (n = 3)",
  "Female", "61.57% (n = 338)", "60.18% (n = 272)", "67.44% (n = 58)",	"72.73% (n = 8)",
  "No degree",	"16.24% (n = 98)", "10.56% (n = 47)",	"39.53% (n = 34)",	"63.54% (n = 7)",
  "GED", "51.58% (n = 279)",	"51.91% (n = 231)",	"52.33% (n = 45)",	"27.27% (n = 3)",
  "Further education", "32.29% (n = 175)", "37.53% (n = 167)", "8.14% (n = 7)",	"9.09% (n = 1)",
  "Apathy",	"0.37 ± 0.63", "0.31 ± 0.59",	"0.65 ± 0.76",	"0.36 ± 0.51",
  "Procrastination",	"28.60 ± 12.00", "27.70 ± 11.30",	"32.10 ± 13.80", "39.00 ± 16.00"
)

# Captions and footnotes
caption <- glue::glue(
  "*Note:* Descriptives for continuous and categorical variables are represented 
  using means ± standard deviations and percentages and frequencies respectively. 
  NC = Normative cognitive function; MCI = Mild cognitive impairment; 
  GED = General Educational Development.")

@@

# Outputting table 
tbl_descriptives |>
  setNames(c(" ", "Full sample (n = 549)", "NC (n = 452)", "MCI (n = 86)", "Dementia (n = 11)")) |>
   tt(notes = list(
    caption, 
    a = list(i = 10, j = 1, text = footnote))
  ) |>
   group_tt(i = list("**Sex**" = 2, "**Education**" = 4)) |>
   style_tt(j = 2:5, align = "c") |>
   style_tt(i = c(2, 5), align = "l") |>
   format_tt(escape = TRUE) |>
   format_tt("notes", markdown = TRUE) |>
   format_tt(j = 1, markdown = TRUE) |>
  style_tt(
     bootstrap_class = "table table-sm"
  )

```

@@

# References

>>>>

This property asserts that the probability of transitioning from state $X_t = i$
to a future state $X\_{t+1} = j$ depends only on the current state $X_t$, and
not on the full history of preceding states. Here, there are three potential
states: normative cognitive function, mild cognitive impairment, and dementia.
We consider dementia to be an absorbing state, i.e. once an individual reaches
this state, they cannot return to the other states in a future time point.

@@

# Table information
tbl_descriptives <- tibble::tribble(
  ~""                 , ~Full              , ~NC                , ~MCI              , ~Dementia        ,
  "Age (years)"       , "69.70 ± 7.58"    , "69.70 ± 7.55"    , "69.60 ± 7.55"   , "70.00 ± 9.89"  ,
  "Male"              , "38.43% (n = 211)" , "39.82% (n =180)"  , "32.56% (n = 28)" , "27.27% (n = 3)" ,
  "Female"            , "61.57% (n = 338)" , "60.18% (n = 272)" , "67.44% (n = 58)" , "72.73% (n = 8)" ,
  "No degree"         , "16.24% (n = 98)"  , "10.56% (n = 47)"  , "39.53% (n = 34)" , "63.54% (n = 7)" ,
  "GED"               , "51.58% (n = 279)" , "51.91% (n = 231)" , "52.33% (n = 45)" , "27.27% (n = 3)" ,
  "Further education" , "32.29% (n = 175)" , "37.53% (n = 167)" , "8.14% (n = 7)"   , "9.09% (n = 1)"  ,
  "Apathy"            , "0.37 ± 0.63"     , "0.31 ± 0.59"     , "0.65 ± 0.76"    , "0.36 ± 0.51"   ,
  "Procrastination"   , "28.60 ± 12.00"   , "27.70 ± 11.30"   , "32.10 ± 13.80"  , "39.00 ± 16.00"
)

# Captions and footnotes
caption <- glue::glue(
  "Note: Descriptives for continuous and categorical variables are represented
  using means ± standard deviations and percentages and frequencies respectively.
  NC = Normative cognitive function; MCI = Mild cognitive impairment;
  GED = General Educational Development.")

@@

# Outputting table
tbl_descriptives |>
  setNames(c(" ", "Full sample (n = 549)", "NC (n = 452)", "MCI (n = 86)", "Dementia (n = 11)")) |>
  tt(notes = list(
    caption,
    a = list(i = 10, j = 1, text = footnote))
  ) |>
   group_tt(i = list("Sex" = 2, "Education" = 4)) |>
   style_tt(j = 2:5, align = "c") |>
   style_tt(i = c(2, 5), align = "l", bold = TRUE) |>
   format_tt(escape = TRUE) |>
   format_tt("notes", markdown = TRUE) |>
   style_tt(
     bootstrap_class = "table table-sm"
     )
```

@@

\newpage

::: {.content-visible when-format="html"}
# References
:::
